Arterial pulsations cannot drive intramural periarterial drainage: Significance for \Ab drainage

### 2017-08-03 paper.tex
added Maureen as author
changed order of authors
modifications for final version

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Diem {et~al.}} %use et al only if is more than 1 author
\def\Authors{Alexandra K. Diem\,$^{1,*}$, Neil W. Bressloff\,$^{1}$, Roxana O. Carare\,$^{2}$, Matthew MacGregor Sharp\,$^{2}$ and Giles Richardson\,$^3$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO16 7QF, UK \\
$^{2}$Clinical Neurosciences, Faculty of Medicine, University of Southampton, Southampton General Hospital, Southampton SO16 6YD, UK\\
$^{3}$Mathematical Sciences, Faculty of Social, Human and Mathematical Sciences, University of Southampton, Southampton SO17 1BJ, UK}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{A.~K.~Diem}
\def\corrAddress{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO16 7QF, UK}
\def\corrEmail{A.K.Diem@soton.ac.uk}

@@

In order to process metabolites, the brain requires a lymphatic system that transports waste products to the lymph nodes, where they are decomposed. While the rest of the body possesses a dedicated network of lymphatic vessels these do not exist in the brain \cite{Weller2010}. Both the pathology of CAA and ample evidence from injection studies suggest that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as intramural periarterial drainage (IPAD) \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}. It is important not to confuse the IPAD pathway with the Virchow-Robin spaces, which are often referred to as perivascular (or ``paravascular'') spaces, possibly artefactual, located between the glia limitans and the arterial wall. Virchow-Robin spaces are virtual spaces and have no role in \Ab drainage. The BM forming the IPAD does not actually constitute a space, but rather a fluid-filled protein matrix. Our understanding of the mechanism that drives perivascular drainage through the BM is still very limited. The evidence indicates that ISF enters the BM at the capillary level and flows towards the lymph nodes in the neck via the BM cerebral arteries. This suggests that the counterintuitive phenomenon of ISF flow occurs in the reverse direction to blood flow \cite{Carare2008}.

@@

Brains were cut into \SI{20}{\micro\metre} coronal sections using a freezing microtome (Leica) and the injection site located by examining on a Zeiss Axioskop 2 fitted with a rhodamine filter. Sections of \SI{400}{\micro\metre} and \SI{800}{\micro\metre} thickness both posterior and anterior to the injection site were selected for immunohistochemistry. Sections were blocked in \SI{15}{\percent} goat serum (Sigma 9023) for \SI{1}{\hour} at room temperature (RT). They were then incubated in primary antibodies diluted in \SI{0.01}{\Molar} phosphate buffered saline \SI{0.1}{\percent} triton x100 (PBSt). Primary antibodies comprised rabbit anti-collagen IV (AbCam, ab6586) 1/400 and mouse anti-smooth muscle actin (SMA) FITC conjugated (Sigma F3777) 1/200. Sections were incubated overnight at \SI{4}{\celsius} in a moist chamber. Following washing 3x \SI{10}{\minute} \SI{0.01}{\Molar} PBS sections were incubated in goat anti-rabbit Alexa Fluor 633 (ThermoFisher Scientific A-21070) 1/200 PBSt.

@@

A.K.D. and G.R. designed and analysed BM model and elasticity analysis, A.K.D., N.W.B and G.R. designed and analysed numerical models, R.O.C and M.M.S. designed and performed mouse experiments, A.K.D., N.W.B, M.M.S., R.O.C. and G.R. wrote the paper.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/pvs_model.pdf}
\textbf{\refstepcounter{figure}\label{fig:pvs_model} Figure \arabic{figure}.}{Geometry of the perivascular drainage pathways model. The length $L$ of a typical artery in the brain can be assumed to be much greater than the width $W$ of the perivascular drainage pathways. Therefore we can approximate the flow as a lubrication model. The pulsation of the capillary is described by the function $h(z,t)$.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/inlet.png}
\textbf{\refstepcounter{figure}\label{fig:inlet} Figure \arabic{figure}.}{Inlet function for MCA simulations obtained from patient-specific measurements used to represent one period of \SI{0.85}{\second} of the pulse wave. The final ten peaks of the velocity time series were averaged in the Fourier space and adjusted such that its range of values aligns with those reported in \cite{Olufsen2002}. The resulting inlet function is smooth and serves as the inlet boundary condition for the MCA simulations using VaMpy \cite{Diem2016a}.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/mca.pdf}
\textbf{\refstepcounter{figure}\label{fig:mca} Figure \arabic{figure}.}{Blood pressure (a), wall displacement (b) and the resulting ISF pressure (c) in the MCA. Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in Table~\ref{tab:simulation}. Wall stiffness is high in small blood vessels, therefore pressure gradients in time are steep. Wall displacement was calculated from the linear elasticity approximation derived in SI~2. The input function $R_i(z,t)$ to the BM model equation \eqref{eq:pvs_model} is obtained by adding the radius at rest $a$ (see geometry depiction in Figure~\ref{fig:pvs_model}).}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/h_dh.png}
\textbf{\refstepcounter{figure}\label{fig:h} Figure \arabic{figure}.}{BM width $h_\textrm{bm}(z,t)$ at several time points during the cardiac cycle (a) and its gradient along the $z$-axis at the same time points (b). Pressure pulse driven variations of the BM width are small compared to the initial width of \SI{200}{\nano\metre} and thus flow through the BM is slow.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/valve_test.png}
\textbf{\refstepcounter{figure}\label{fig:valve_test} Figure \arabic{figure}.}{ISF flux (positive values in the direction of arterial flow) averaged over one cycle as a function of BM position $\eta$ in the arterial wall. Here the BM is at $r = r_0 + \eta h$ such that $\eta = 0$ represents a BM immediately adjacent to the lumen and $\eta = 1$ represents a BM on the outer wall of the artery. Note that reverse flow is only achieved by ratios of $K_0/K_1 < \SI{2.74e-2}{}$. The fastest net reverse drainage shown here (for $K_0/K_1 = \SI{2.68e-2}{}$ and on the arterial lumen $\eta = 0$) is \SI{-2.64e-5}{\cubic\micro\metre\per\second}.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/confocal.jpg}
\textbf{\refstepcounter{figure}\label{fig:confocal} Figure \arabic{figure}.}{Composite tile scan of confocal images of the distribution of intracerebrally injected \Ab, in relation to collagen IV and SMA. Immunostaining \Ab: red, SMA: green, collagen IV: blue. Arrows are pointing to arteries with \Ab in the intramural perivascular drainage pathways.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/graph.png}
\textbf{\refstepcounter{figure}\label{fig:graph} Figure \arabic{figure}.}{Graphs showing the number of arteries, veins and capillaries with \Ab in their walls in atenolol treated mice compared to control mice. SEM: standard error of mean.}
\end{figure}

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Diem {et~al.}} %use et al only if is more than 1 author
\def\Authors{Alexandra K. Diem\,$^{1,*}$, Matthew MacGregor Sharp\,$^{2}$, Maureen Gatherer\,$^{2}$, Neil W. Bressloff\,$^{1}$, Roxana O. Carare\,$^{2}$ and Giles Richardson\,$^3$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO16 7QF, UK \\
$^{2}$Clinical Neurosciences, Faculty of Medicine, University of Southampton, Southampton General Hospital, Southampton SO16 6YD, UK\\
$^{3}$Mathematical Sciences, Faculty of Social, Human and Mathematical Sciences, University of Southampton, Southampton SO17 1BJ, UK}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{A.~K.~Diem}
\def\corrAddress{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO16 7QF, UK}
\def\corrEmail{A.K.Diem@soton.ac.uk}

@@

In order to process metabolites, the brain requires a lymphatic system that transports waste products to the lymph nodes. While the rest of the body possesses a dedicated network of lymphatic vessels these do not exist in the brain \cite{Weller2010}. Both the pathology of CAA and ample evidence from injection studies suggest that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as intramural periarterial drainage (IPAD) \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}. It is important not to confuse the IPAD pathway with the Virchow-Robin spaces, which are often referred to as perivascular (or ``paravascular'') spaces, possibly artefactual, located between the glia limitans and the arterial wall. Virchow-Robin spaces are compartments occupied by the pial-glial basement membranes spaces and have no role in \Ab drainage from the extracellular spaces. The BM forming the IPAD does not actually constitute a space, but rather a fluid-filled protein matrix. Our understanding of the mechanism that drives perivascular drainage through the BM is still very limited. The evidence indicates that ISF enters the BM at the capillary level and flows towards the lymph nodes in the neck via the BM cerebral arteries. This suggests that the counterintuitive phenomenon of ISF flow occurs in the reverse direction to blood flow \cite{Carare2008}.

@@

Brains were cut into \SI{20}{\micro\metre} coronal sections using a freezing microtome (Leica) and the injection site located by examining on a Zeiss Axioskop 2 fitted with a rhodamine filter. Sections \SI{400}{\micro\metre} and \SI{800}{\micro\metre} both posterior and anterior to the injection site were selected for immunohistochemistry. Sections were blocked in \SI{15}{\percent} goat serum (Sigma 9023) for \SI{1}{\hour} at room temperature (RT). They were then incubated in primary antibodies diluted in \SI{0.01}{\Molar} phosphate buffered saline \SI{0.1}{\percent} triton x100 (PBSt). Primary antibodies comprised rabbit anti-collagen IV (AbCam, ab6586) 1/400 and mouse anti-smooth muscle actin (SMA) FITC conjugated (Sigma F3777) 1/200. Sections were incubated overnight at \SI{4}{\celsius} in a moist chamber. Following washing 3x \SI{10}{\minute} \SI{0.01}{\Molar} PBS sections were incubated in goat anti-rabbit Alexa Fluor 633 (ThermoFisher Scientific A-21070) 1/200 PBSt.

@@

A.K.D. and G.R. designed and analysed BM model and elasticity analysis, A.K.D., N.W.B and G.R. designed and analysed numerical models, R.O.C and M.M.S. designed and performed mouse experiments, M.G. analysed brain sections and performed image analysis, A.K.D., N.W.B, M.M.S., R.O.C. and G.R. wrote the paper.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig01.jpg}
\textbf{\refstepcounter{figure}\label{fig:pvs_model} Figure \arabic{figure}.}{Geometry of the perivascular drainage pathways model. The length $L$ of a typical artery in the brain can be assumed to be much greater than the width $W$ of the perivascular drainage pathways. Therefore we can approximate the flow as a lubrication model. The pulsation of the capillary is described by the function $h(z,t)$.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig02.jpg}
\textbf{\refstepcounter{figure}\label{fig:inlet} Figure \arabic{figure}.}{Inlet function for MCA simulations obtained from patient-specific measurements used to represent one period of \SI{0.85}{\second} of the pulse wave. The final ten peaks of the velocity time series were averaged in the Fourier space and adjusted such that its range of values aligns with those reported in \cite{Olufsen2002}. The resulting inlet function is smooth and serves as the inlet boundary condition for the MCA simulations using VaMpy \cite{Diem2016a}.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig03.jpg}
\textbf{\refstepcounter{figure}\label{fig:mca} Figure \arabic{figure}.}{Blood pressure (a), wall displacement (b) and the resulting ISF pressure (c) in the MCA. Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in Table~\ref{tab:simulation}. Wall stiffness is high in small blood vessels, therefore pressure gradients in time are steep. Wall displacement was calculated from the linear elasticity approximation derived in SI~2. The input function $R_i(z,t)$ to the BM model equation \eqref{eq:pvs_model} is obtained by adding the radius at rest $a$ (see geometry depiction in Figure~\ref{fig:pvs_model}).}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig04.jpg}
\textbf{\refstepcounter{figure}\label{fig:h} Figure \arabic{figure}.}{BM width $h_\textrm{bm}(z,t)$ at several time points during the cardiac cycle (a) and its gradient along the $z$-axis at the same time points (b). Pressure pulse driven variations of the BM width are small compared to the initial width of \SI{200}{\nano\metre} and thus flow through the BM is slow.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig05.jpg}
\textbf{\refstepcounter{figure}\label{fig:valve_test} Figure \arabic{figure}.}{ISF flux (positive values in the direction of arterial flow) averaged over one cycle as a function of BM position $\eta$ in the arterial wall. Here the BM is at $r = r_0 + \eta h$ such that $\eta = 0$ represents a BM immediately adjacent to the lumen and $\eta = 1$ represents a BM on the outer wall of the artery. Note that reverse flow is only achieved by ratios of $K_0/K_1 < \SI{2.74e-2}{}$. The fastest net reverse drainage shown here (for $K_0/K_1 = \SI{2.68e-2}{}$ and on the arterial lumen $\eta = 0$) is \SI{-2.64e-5}{\cubic\micro\metre\per\second}.}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig06.jpg}
\textbf{\refstepcounter{figure}\label{fig:confocal} Figure \arabic{figure}.}{Composite tile scan of confocal images of the distribution of intracerebrally injected \Ab, in relation to collagen IV and SMA. The image was taken at \SI{400}{\micro\m} anterior to the injection site in an atenolol (Ate) injected mouse, to avoid any artefacts induced by the injection tract, relying instead on the pattern of distribution of the fluorescently injected Abeta away from the direct injection. Immunostaining \Ab: red, SMA: green, collagen IV: blue. Arrows are pointing to arteries with \Ab in the intramural perivascular drainage pathways. Scale bars = \SI{400}{\micro\m}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10cm]{figures/fig07.jpg}
\textbf{\refstepcounter{figure}\label{fig:graph} Figure \arabic{figure}.}{Graphs showing the number of arteries, veins and capillaries with \Ab in their walls in atenolol treated mice compared to control mice. SEM: standard error of mean.}
\end{figure}
